Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}

### 2023-05-09 Papers/Paper CASC format/tmp/FCpaper.tex
...

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{ifthen}
\usepackage{mmacells}
\usepackage[invisible]{MLaTeX}

@@

\begin{document}
\title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
\author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
	Friedemann Krannich\inst{1}\orcidID{1111-2222-3333-4444} \and
	Bashayer Majrashi\inst{1}\orcidID{0009-0008-3677-8064 } \and
	Ricardo de Lima Ribeiro\inst{1}\orcidID{0000-0002-2707-2210}}
\authorrunning{D.A. Gomes et al.}
\institute{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia\
	\email{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeiro}@kaust.edu.sa}}
\maketitle
\begin{abstract}
	Many partial differential equations (PDEs) admit conserved quantities like mass or energy. These quantities are often essential for establishing well-posedness results. When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme. Such conservation may ensure the stability of the numerical scheme. We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme. Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. Additionally, we introduce an algorithm for finding conserved quantities, given a scheme. We illustrate our algorithm with several finite-difference schemes. Our approach incorporates parametric G\"obner bases to handle parameters, ensuring accurate computation of conserved quantities.

@@

{\color{red} $\Phi$ and $\Psi$ need domains and ranges, also explain that divergence is only in $x$}

Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
\begin{gather*}
	\frac{\partial }{\partial t}\Phi[u]+\div \Psi[u]=0
\end{gather*}
holding for all solutions $u$ of the considered PDE,
that induces the conserved quantity $\int \Phi[u]\ dx$ as
\begin{gather*}
	\frac{d}{dt}\int \Phi[u]\ dx=\int D_t\Phi[u]\ dx=-\int D_x\Psi[u]\ dx=0
\end{gather*}
assuming periodic boundary conditions in $x$.

@@

In this paper, we present an algorithm that verifies if a quantity is conserved over time under a given scheme. Our approach diverges from previous methods, as we do not attempt to construct discrete $\Phi$ and $\Psi$ using the scheme, but rather check if a given discrete $\Phi$ is conserved over time under the scheme. Typically, conserved quantities involve integral expressions or their discrete analogs, such as sums. Gomes et al. proposed algorithms for simplifying sums \cite{gomes20} and developed algorithms for detecting conserved quantities in PDEs and semi-discretized schemes \cite{gomes20}. We review these techniques in Section \ref{assp}.
To generalize these methods for cases without summing over all arguments of the involved functions, we introduce the discrete partial variational derivative (Section \ref{pvard}). Our main contribution is an algorithm for verifying the conservation of a quantity under a numerical scheme. Gerdt \cite{gerdt12} demonstrated that the quantity is conserved if its discrete time derivative belongs to the difference ideal generated by the scheme. However, some quantities may add only to a constant and be trivially preserved without being part of the difference ideal. Furthermore, Gerdt's algorithm may not always terminate, as the Gr{\"o}bner basis for the difference ideal might be infinite.
We address these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) having a finite Gr{\"o}bner basis (Section \ref{algodetec}). Our algorithm is applicable to schemes that are explicit and implicit in time, can handle schemes with multiple higher-order time and space derivatives, multiple space dimensions, systems of equations, and schemes with parameters.
As the process of finding conserved quantities involves polynomial division, special care must be taken to ensure that the division is performed correctly with respect to the parameters. We employ parametric Gr{\"o}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section \ref{pgb}
 We have implemented this algorithm as a part of a {\sc Mathematica} package \cite{wolfram21}\footnote{The code is available upon request. For access, contact ricardo.ribeiro@kaust.edu.sa.}. Our examples include Burgers equation and a PDE system related to mean-field games (Section \ref{eaa}). In these examples, 
 we show that our code detects conserved quantities and can be used to select suitable schemes for time-implicit and time-explicit discretizations.

@@

In this section, we introduce the concept of discrete torus and discrete variational derivatives, and discuss their properties and applications in simplifying sums. 
These definitions are essential for the analysis of methods presented in the rest of the paper.
Here, we 
discuss the connection between discrete functionals and discrete variational derivatives and present examples that illustrate how to simplify sums using this approach.
We follow the ideas in \cite{gomes20}.
In this paper, subscripts denote coordinate or tuple indices, and superscripts denote sequence indices.
To avoid boundary terms, we work with periodic functions in $\Zz^d$.

When discretizing PDEs using finite difference schemes, we need to work with functions defined on discrete sets. For instance, periodic functions in $\Rr^d$ can be replaced by functions defined on the discrete torus, which will be introduced in the following discussion.

@@

	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is the set $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space

@@

		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\}.

@@

Functions in $\mathcal{P}(\Ii,\Rr^m)$ extend periodically to $\Zz^d$.

@@

	The space of functionals $\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$ (not necessarily linear) on $\mathcal{P}(\Ii,\Rr^m)$ is

@@

		\mathcal{F}(\mathcal{P}(\Ii,\Rr^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\Ii,\Rr^m)\to \Rr\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\Ii}F_n[u],\\

@@

		\text{ and }\{e^1,\ldots,e^k\}\subset\Zz^d,\ \text{G is smooth}

@@

In the previous definition, $e^1,\ldots,e^k$ are not necessarily unit vectors and may vary between functionals.

@@

	Let $d=m=1$ and
	$	\mathcal{J}[u]:=\sum_{n}(u(n+1)-u(n))^2$.

@@

Let $u,v\in\mathcal{P}(\Ii,\Rr^m)$, $\epsilon 
\in\Rr$ and $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. Define

@@

In \eqref{periodicity}, we shift the indices because of the periodicity of $u$ and $v$,

@@

	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
	\notag	=\sum_{n\in\Ii}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
	\notag	=\sum_{n\in\Ii}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\

@@

	\label{periodicity}	=\sum_{n\in\Ii}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\

@@

	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\

@@

	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).	

@@

The operator $\mathcal{V}(\mathcal{J})$ can be viewed as a representation of the derivative $D\mathcal{J}u$ with respect to the $L^2$ inner product $\langle w, v\rangle=\sum_{n\in \Ii} w_n v_n$, leading to the following definition.

@@

	$\mathcal{V}(\mathcal{J})$ is the \emph{discrete variational derivative} of $\mathcal{J}$.

@@

		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
		=\sum_{n\in\Ii}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
		=\sum_{n\in\Ii}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
		=\sum_{n\in\Ii}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
		=\sum_{n\in\Ii}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).

@@

Thus, $\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)$.

@@

\begin{theorem}\label{theo1}
	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. If $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
	$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
	for all $u\in\mathcal{P}(\Ii,\Rr^m)$.
	Conversely,
	$\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
	for all $u\in\mathcal{P}(\Ii,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{theorem}
\begin{proof}
	We have
	\begin{gather*}
		(\mathcal{J}-\tilde{\mathcal{J}})[u]=(\mathcal{J}-\tilde{\mathcal{J}})[u]-(\mathcal{J}-\tilde{\mathcal{J}})[u_0]\\
		=\int_{0}^{1}\frac{d}{d\lambda}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)]\ d\lambda\\
		=\int_{0}^{1}\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)](u-u_0)\ d\lambda=0
	\end{gather*}
	and hence $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{proof}
We use Theorem \ref{theo1} to determine whether different sums represent the same quantity.
\begin{example}
	Let $d=m=1$ and consider the functionals
	\begin{gather*}
		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1)\text{ and }
		\tilde{\mathcal{J}}[u]:=\sum_{n\in\Ii}u(n-2)u(n-1).
	\end{gather*}
	It is clear that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
	\begin{gather*}
		D(\mathcal{J}-\tilde{\mathcal{J}})[u](v)=\\
		=\sum_{n\in\Ii}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
	\end{gather*}
	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals represent the same quantity.
\end{example}

\begin{example}
	Let $d=2$ and $m=1$. It may not be obvious, that
	\begin{gather*}
		\sum_{n_1,n_2}u(n_1,n_2-2)^4 - 3 u(n_1,n_2-2)^3 u(n_1+1,n_2-2) \\+ 
		u(n_1,n_2) u(n_1+1,n_2)^3 + 3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2\\ -
		u(n_1,n_2-2) u(n_1+1,n_2-2)^3
		- u(n_1, n_2)^4 +\\
		3 u(n_1,n_2)^3 u(n_1+1,n_2)- 3 u(n_1,n_2)^2 u(n_1+1,n_2)\\
		-3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2+ 3 u(n_1,n_2)^2 u(n_1+1,n_2)=0.
	\end{gather*}
	We confirm this by computing the discrete variational derivative and noticing that the expression is $0$ for $u=0$.
\end{example}

@@

In this section, we introduce the concept of the discrete partial variational derivative, a generalization of the discrete variational derivative for cases where one or several arguments of the indices are kept constant. This generalization provides a more flexible approach to handling various problems in conservation laws for finite-difference schemes, addressing challenges and limitations of the traditional discrete variational derivative. The main result of this section is Theorem \ref{theo2}, which gives a criterion for equality of two functionals based on their discrete partial variational derivatives."

Let $n=(n_1,\ldots,n_d)$ be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$. We call $n$ the space variables. Let $l\in\Nn$ and call $t:=(t_1,\ldots,t_l)\in\Nn^l$ the time variables. For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when $l=1$.
The analog to Definition \ref{d1} is the following. 
\begin{definition}
	Let $\mathcal{P}(\Ii\times\Nn^l,\Rr^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\Ii,\Rr^m)\ \forall t\in\Nn^l \}$.
	Functions in $\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ extend to $\Zz^d\times\Nn^l$ through  periodicity in the space variables.
	Let $\mathcal{D}(\Nn^l):=\{u\ |\ u:\Nn^l\rightarrow\Rr\}$.
\end{definition}
We can regard $\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ as the set of functions depending in space and time and $\mathcal{D}(\Nn^l)$ as the functions depending only on time.

Next, we examine functions that involve summation only over spatial variables, extending the concept presented in Definition \ref{d2}.
\begin{definition}
	\label{quantityfunc2}
	The space of functions $\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$ is
	\begin{gather*}
		\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u](t)=\sum_{n\in\Ii}F_{(n,t)}[u],\\
		\text{where } F_{(n,t)}[u]=G(u((n,t)+e^1),u((n,t)+e^2),\ldots,u((n,t)+e^k)),\\
		\text{and }
		\{e^1,\ldots,e^k\}\subset\Zz^{d+l},\text{G is a polynomial in k variables}\Bigg\}.
	\end{gather*}
	Here, $(n,t):=(n_1,\ldots,n_d,t_1,\ldots,t_l)\in\Ii\times\Nn^l$.
\end{definition}

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[latin1]{inputenc}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{ifthen}
\usepackage{mmacells}

@@

\usepackage{color}\begin{document}

\title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
\author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
	Friedemann Krannich\inst{1}\orcidID{1111-2222-3333-4444} \and
	Bashayer Majrashi\inst{1}\orcidID{0009-0008-3677-8064 } \and
	Ricardo de Lima Ribeiro\inst{1}\orcidID{0000-0002-2707-2210}}
\authorrunning{D.A. Gomes et al.}
\institute{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia\
	\email{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeiro}@kaust.edu.sa}}
\maketitle
\begin{abstract}
	Many partial differential equations (PDEs) admit conserved quantities like mass or energy. These quantities are often essential for establishing well-posedness results. When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme. Such conservation may ensure the stability of the numerical scheme. We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme. Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. Additionally, we introduce an algorithm for finding conserved quantities, given a scheme. We illustrate our algorithm with several finite-difference schemes. Our approach incorporates parametric G\"obner bases to handle parameters, ensuring accurate computation of conserved quantities.

@@

{\color{red} \(\Phi\)
{\color{blue} New variable(s): \(\Phi\),  }
 and \(\Psi\)
{\color{blue} New variable(s): \(\Psi\),  }
 need domains and ranges, also explain that divergence is only in \(x\)
{\color{blue} New variable(s): \(x\),  }
}

Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
\begin{gather*}
	\frac{\partial }{\partial t}\Phi[u]+\operatorname{div} \Psi[u]=0
\end{gather*}
holding for all solutions \(u\)
{\color{blue} New variable(s): \(u\),  }
 of the considered PDE,
that induces the conserved quantity \(\int \Phi[u]\ dx\)
{\color{blue} New variable(s): \(d\),  }
 as
\begin{gather*}
	\frac{d}{dt}\int \Phi[u]\ dx=\int D_t\Phi[u]\ dx=-\int D_x\Psi[u]\ dx=0
\end{gather*}
assuming periodic boundary conditions in \(x\).

@@

In this paper, we present an algorithm that verifies if a quantity is conserved over time under a given scheme. Our approach diverges from previous methods, as we do not attempt to construct discrete \(\Phi\) and \(\Psi\) using the scheme, but rather check if a given discrete \(\Phi\) is conserved over time under the scheme. Typically, conserved quantities involve integral expressions or their discrete analogs, such as sums. Gomes et al. proposed algorithms for simplifying sums \cite{gomes20} and developed algorithms for detecting conserved quantities in PDEs and semi-discretized schemes \cite{gomes20}. We review these techniques in Section \ref{assp}
 {\color{red} Forward or undefined reference: \begin{verbatim}\ref{assp}\end{verbatim}.}
.
To generalize these methods for cases without summing over all arguments of the involved functions, we introduce the discrete partial variational derivative (Section \ref{pvard}
 {\color{red} Forward or undefined reference: \begin{verbatim}\ref{pvard}\end{verbatim}.}
). Our main contribution is an algorithm for verifying the conservation of a quantity under a numerical scheme. Gerdt \cite{gerdt12} demonstrated that the quantity is conserved if its discrete time derivative belongs to the difference ideal generated by the scheme. However, some quantities may add only to a constant and be trivially preserved without being part of the difference ideal. Furthermore, Gerdt's algorithm may not always terminate, as the Gr{\"o}bner basis for the difference ideal might be infinite.
We address these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) having a finite Gr{\"o}bner basis (Section \ref{algodetec}
 {\color{red} Forward or undefined reference: \begin{verbatim}\ref{algodetec}\end{verbatim}.}
). Our algorithm is applicable to schemes that are explicit and implicit in time, can handle schemes with multiple higher-order time and space derivatives, multiple space dimensions, systems of equations, and schemes with parameters.
As the process of finding conserved quantities involves polynomial division, special care must be taken to ensure that the division is performed correctly with respect to the parameters. We employ parametric Gr{\"o}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section \ref{pgb}
 {\color{red} Forward or undefined reference: \begin{verbatim}\ref{pgb}\end{verbatim}.}

 We have implemented this algorithm as a part of a {\sc Mathematica} package \cite{wolfram21}\footnote{The code is available upon request. For access, contact ricardo.ribeiro@kaust.edu.sa.}. Our examples include Burgers equation and a PDE system related to mean-field games (Section \ref{eaa}
 {\color{red} Forward or undefined reference: \begin{verbatim}\ref{eaa}\end{verbatim}.}
). In these examples, 
 we show that our code detects conserved quantities and can be used to select suitable schemes for time-implicit and time-explicit discretizations.

@@

In this section, we introduce the concept of discrete torus and discrete variational derivatives, and discuss their properties and applications in simplifying sums. 
These definitions are essential for the analysis of methods presented in the rest of the paper.
Here, we 
discuss the connection between discrete functionals and discrete variational derivatives and present examples that illustrate how to simplify sums using this approach.
We follow the ideas in \cite{gomes20}.
In this paper, subscripts denote coordinate or tuple indices, and superscripts denote sequence indices.
To avoid boundary terms, we work with periodic functions in \({\mathbb{Z}}^d\)
{\color{blue} New variable(s): \(Z\),  }
.

When discretizing PDEs using finite difference schemes, we need to work with functions defined on discrete sets. For instance, periodic functions in \(\mathbb{R}^d\)
{\color{blue} New variable(s): \(R\),  }
 can be replaced by functions defined on the discrete torus, which will be introduced in the following discussion.

@@

	Let \(N\)
{\color{blue} New variable(s): \(N\),  }
, \(d\) and \(m\)
{\color{blue} New variable(s): \(m\),  }
 be positive integers. The \emph{discrete torus} is the set \(\mathcal{I}:=\{0,\ldots,N-1\}^d\subset {\mathbb{Z}}^d\). Define the space

@@

		\mathcal{P}(\mathcal{I},\mathbb{R}^m):=\{u\ |\ u:\mathcal{I}\rightarrow\mathbb{R}^m\}.

@@

Functions in \(\mathcal{P}(\mathcal{I},\mathbb{R}^m)\)
{\color{blue} New variable(s): \(P\),  }
 extend periodically to \({\mathbb{Z}}^d\).

@@

	The space of functionals \(\mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m))\)
{\color{blue} New variable(s): \(F\),  }
 (not necessarily linear) on \(\mathcal{P}(\mathcal{I},\mathbb{R}^m)\) is

@@

		\mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\mathcal{I},\mathbb{R}^m)\to \mathbb{R}\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\mathcal{I}}F_n[u],\\

@@

		\text{ and }\{e^1,\ldots,e^k\}\subset{\mathbb{Z}}^d,\ \text{G is smooth}

@@

In the previous definition, \(e^1,\ldots,e^k\)
{\color{blue} New variable(s): \(k\),  }
 are not necessarily unit vectors and may vary between functionals.

@@

	Let \(d=m=1\) and
	\(	\mathcal{J}[u]:=\sum_{n}(u(n+1)-u(n))^2\)
{\color{blue} New variable(s): \(J\), \(n\),  }
.

@@

Let \(u,v\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\)
{\color{blue} New variable(s): \(v\),  }
, \(\epsilon 
\in\mathbb{R}\)
{\color{blue} New variable(s): \(\epsilon\),  }
 and \(\mathcal{J}\in\mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m))\). Define

@@

In \eqref{periodicity}
 {\color{red} Forward or undefined reference: \begin{verbatim}\eqref{periodicity}\end{verbatim}.}
, we shift the indices because of the periodicity of \(u\) and \(v\),

@@

	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\mathcal{I}}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
	\notag	=\sum_{n\in\mathcal{I}}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
	\notag	=\sum_{n\in\mathcal{I}}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\

@@

	\label{periodicity}	=\sum_{n\in\mathcal{I}}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\

@@

	\notag	=\sum_{n\in\mathcal{I}}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\

@@

	\notag	=:\sum_{n\in\mathcal{I}}\mathcal{V}(\mathcal{J})[u](n)v(n).	

@@

The operator \(\mathcal{V}(\mathcal{J})\)
{\color{blue} New variable(s): \(V\),  }
 can be viewed as a representation of the derivative \(D\mathcal{J}u\)
{\color{blue} New variable(s): \(D\),  }
 with respect to the \(L^2\)
{\color{blue} New variable(s): \(L\),  }
 inner product \(\langle w, v\rangle=\sum_{n\in \mathcal{I}} w_n v_n\)
{\color{blue} New variable(s): \(w\), \(v_n\), \(w_n\),  }
, leading to the following definition.

@@

	\(\mathcal{V}(\mathcal{J})\) is the \emph{discrete variational derivative} of \(\mathcal{J}\).

@@

		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\mathcal{I}}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
		=\sum_{n\in\mathcal{I}}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
		=\sum_{n\in\mathcal{I}}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
		=\sum_{n\in\mathcal{I}}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
		=\sum_{n\in\mathcal{I}}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
		=\sum_{n\in\mathcal{I}}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
		=\sum_{n\in\mathcal{I}}\mathcal{V}(\mathcal{J})[u](n)v(n).

@@

Thus, \(\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)\).

@@

\begin{theorem}\label{theo1}
	Let \(\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m))\). If \(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0\) for all \(v\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\) and if there exists \(u_0\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\)
{\color{blue} New variable(s): \(u_0\),  }
 such that \(\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]\), then
	\(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\)
	for all \(u\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\).
	Conversely,
	\(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0\)
	for all \(u\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\), if \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
\end{theorem}
\begin{proof}
	We have
	\begin{gather*}
		(\mathcal{J}-\tilde{\mathcal{J}})[u]=(\mathcal{J}-\tilde{\mathcal{J}})[u]-(\mathcal{J}-\tilde{\mathcal{J}})[u_0]\\
		=\int_{0}^{1}\frac{d}{d\lambda}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)]\ d\lambda\\
		=\int_{0}^{1}\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)](u-u_0)\ d\lambda=0
	\end{gather*}
	and hence \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
\end{proof}
We use Theorem \ref{theo1} to determine whether different sums represent the same quantity.
\begin{example}
	Let \(d=m=1\) and consider the functionals
	\begin{gather*}
		\mathcal{J}[u]:=\sum_{n\in\mathcal{I}}u(n)u(n+1)\text{ and }
		\tilde{\mathcal{J}}[u]:=\sum_{n\in\mathcal{I}}u(n-2)u(n-1).
	\end{gather*}
	It is clear that \(\mathcal{J}\) and \(\tilde{\mathcal{J}}\) represent the same quantity (by shifting \(n\)). We confirm this, using the discrete variational derivative:
	\begin{gather*}
		D(\mathcal{J}-\tilde{\mathcal{J}})[u](v)=\\
		=\sum_{n\in\mathcal{I}}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
	\end{gather*}
	Hence, \(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0\) and \(\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]\). Therefore, both functionals represent the same quantity.
\end{example}

\begin{example}
	Let \(d=2\) and \(m=1\). It may not be obvious, that
	\begin{gather*}
		\sum_{n_1,n_2}u(n_1,n_2-2)^4 - 3 u(n_1,n_2-2)^3 u(n_1+1,n_2-2) \\+ 
		u(n_1,n_2) u(n_1+1,n_2)^3 + 3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2\\ -
		u(n_1,n_2-2) u(n_1+1,n_2-2)^3
		- u(n_1, n_2)^4 +\\
		3 u(n_1,n_2)^3 u(n_1+1,n_2)- 3 u(n_1,n_2)^2 u(n_1+1,n_2)\\
		-3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2+ 3 u(n_1,n_2)^2 u(n_1+1,n_2)=0.
	\end{gather*}
	We confirm this by computing the discrete variational derivative and noticing that the expression is \(0\)
{\color{blue} New variable(s): \(\text{Null}\),  }
 for \(u=0\).
\end{example}

@@

In this section, we introduce the concept of the discrete partial variational derivative, a generalization of the discrete variational derivative for cases where one or several arguments of the indices are kept constant. This generalization provides a more flexible approach to handling various problems in conservation laws for finite-difference schemes, addressing challenges and limitations of the traditional discrete variational derivative. The main result of this section is Theorem \ref{theo2}
 {\color{red} Forward or undefined reference: \begin{verbatim}\ref{theo2}\end{verbatim}.}
, which gives a criterion for equality of two functionals based on their discrete partial variational derivatives."

Let \(n=(n_1,\ldots,n_d)\)
{\color{blue} New variable(s): \(n_1\), \(n_d\),  }
 be the variables for the functions in \(\mathcal{P}(\mathcal{I},\mathbb{R}^m)\). We call \(n\) the space variables. Let \(l\in\mathbb{N}\)
{\color{blue} New variable(s): \(l\),  }
 and call \(t:=(t_1,\ldots,t_l)\in\mathbb{N}^l\)
{\color{blue} New variable(s): \(t\), \(t_1\), \(t_l\),  }
 the time variables. For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when \(l=1\).
The analog to Definition \ref{d1} is the following. 
\begin{definition}
	Let {\color{red} \(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\ \forall t\in\mathbb{N}^l \}\)

  Equation errors and warnings:

 Unbalanced $|\hdots|$}

.
	Functions in \(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\) extend to \({\mathbb{Z}}^d\times\mathbb{N}^l\) through  periodicity in the space variables.
	Let {\color{red} \(\mathcal{D}(\mathbb{N}^l):=\{u\ |\ u:\mathbb{N}^l\rightarrow\mathbb{R}\}\)

  Equation errors and warnings:

 Unbalanced $|\hdots|$}

.
\end{definition}
We can regard \(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\) as the set of functions depending in space and time and \(\mathcal{D}(\mathbb{N}^l)\) as the functions depending only on time.
